---
title: "Assignment 1 - Linear Programming"
output:
  pdf_document:
    toc: yes
    toc_depth: 5
  beamer_presentation:
    colortheme: lily
    fig_caption: no
    fig_height: 6
    fig_width: 7
    fonttheme: structurebold
    theme: Hannover
    toc: yes
  prettydoc::html_pretty:
    df_print: paged
    highlight: vignette
    theme: architect
    toc: yes
    toc_depth: 5
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  slidy_presentation:
    highlight: default
  ioslides_presentation:
    css:
    - css/fonts.css
    - css/custom.css
    - css/title-slide.css
    - css/slide-background.css
    includes:
      before_body: html/title.html
    toc: yes
    transition: default
    widescreen: yes
course: Decision Models
---

```{r setup, include=FALSE}
# Use 'verbatim = TRUE' as chunk option to show chunk code as is
require(knitr)
hook_source_def = knit_hooks$get('source')
knit_hooks$set(source = function(x, options){
  if (!is.null(options$verbatim) && options$verbatim){
    opts = gsub(",\\s*verbatim\\s*=\\s*TRUE\\s*", "", options$params.src)
    bef = sprintf('\n\n    ```{r %s}\n', opts, "\n")
    stringr::str_c(bef, paste(knitr:::indent_block(x, "    "), collapse = '\n'), "\n    ```\n")
  } else {
     hook_source_def(x, options)
  }
})
```

## The problem
A trading company is looking for a way to maximize profit per transportation of their goods. The company has a train available with 3 wagons. 

When stocking the wagons they can choose among 4 types of cargo, each with its own specifications. 
How much of each cargo type should be loaded on which wagon in order to maximize profit?

### More data
|TRAIN WAGON $j$	|WEIGHT CAPACITY (TONNE) $w_j$	|VOLUME CAPACITY ($m^2$) $s_j$|
|-----------------|-------------------------------|-----------------------------|
|(wag) 1 	        |10	                            |5000                         |
|(wag) 2	        |8	                            |4000                         |
|(wag) 3	        |12	                            |8000                         |
  



|CARGO TYPE	$i$ |AVAILABLE (TONNE) $a_i$	|VOLUME ($m^2$) $v_i$	|PROFIT (PER TONNE) $p_i$ |
|---------------|-------------------------|---------------------|-------------------------|
|(cg) 1	        |18	                      |400	                |2000                     |
|(cg) 2	        |10	                      |300	                |2500                     |
|(cg) 3	        |5	                      |200	                |5000                     |
|(cg) 4	        |20	                      |500	                |3500                     |


### The decision variables

Define the decision variables for the problem described above.

### The objective function

Define the objective function for the problem described above.

### The constraints

Define the constraints for the problem described above.

## Building the model

```{r cars}
#LP_SOLVE
if(require(lpSolveAPI)==FALSE) install.packages("lpSolveAPI")
```
```{r }
library(lpSolveAPI)
model=make.lp(0,12)
lp.control(model, sense="max")
set.objfn(model, obj=c(2000, 2500, 5000, 3500, 2000, 2500, 5000, 3500, 2000, 2500, 5000, 3500))

add.constraint(model, xt=c(1, 1, 1, 1), type="<=", rhs=10, indices=c(1:4))
add.constraint(model, xt=c(1, 1, 1, 1), type="<=", rhs=8, indices=c(5:8))
add.constraint(model, xt=c(1, 1, 1, 1), type="<=", rhs=12, indices=c(9:12))

add.constraint(model, xt=c(400, 300, 200, 500), type="<=", rhs=5000, indices=c(1:4))
add.constraint(model, xt=c(400, 300, 200, 500), type="<=", rhs=4000, indices=c(5:8))
add.constraint(model, xt=c(400, 300, 200, 500), type="<=", rhs=8000, indices=c(9:12))

add.constraint(model, xt=c(1, 1, 1), type="<=", rhs=18, indices=c(1, 5, 9))
add.constraint(model, xt=c(1, 1, 1), type="<=", rhs=10, indices=c(2, 6, 10))
add.constraint(model, xt=c(1, 1, 1), type="<=", rhs=5, indices=c(3, 7, 11))
add.constraint(model, xt=c(1, 1, 1), type="<=", rhs=20, indices=c(4, 8, 12))

set.bounds(model, lower = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)) #vincolo di non negativit? 

solve(model)
get.variables(model)
get.objective(model)
get.constraints(model)
get.primal.solution(model)
get.basis(lprec=model, nonbasic=FALSE)
get.dual.solution(model)

```
## Sensitivity analysis

Perform the sensitivity analysis for the model solved.

```{r cars}
library(dplyr)
library(tidyr)

get.sensitivity.rhs(model)
get.sensitivity.obj(model)

printSensitivityRHS <- function(model){
  options(scipen=999)
  arg.rhs =get.sensitivity.rhs(model)
  numRows <- length(arg.rhs$duals)
  symb <- c()
  for (i in c(1:numRows)) symb[i] <- paste("M", i, sep = "" )
  rhs <- data.frame(rhs = symb,arg.rhs)
  rhs<-rhs %>%
    mutate(dualsfrom=replace(dualsfrom, dualsfrom < -1.0e4, "-inf")) %>%
    mutate(dualstill=replace(dualstill, dualstill > 1.0e4, "inf")) %>%
    unite(col = "Sensitivity",dualsfrom,rhs,dualstill, sep = " <= ", remove = FALSE) %>% 
    select(c("rhs","Sensitivity"))
  colnames(rhs)[1]<-c('Rhs')
  print(rhs)
}

printSensitivityObj <- function(model){
  options(scipen=999)
  arg.obj = get.sensitivity.obj(model)
  numRows <- length(arg.obj$objfrom)
  symb <- c()
  for (i in c(1:numRows)) symb[i] <- paste("N", i, sep = "" )
  obj <- data.frame(Objs = symb, arg.obj)
  obj<-obj %>%
    mutate(objfrom=replace(objfrom, objfrom < -1.0e4, "-inf")) %>%
    mutate(objtill=replace(objtill, objtill > 1.0e4, "inf")) %>%
    unite(col = "Sensitivity",objfrom, Objs, objtill,sep = " <= ", remove = FALSE) %>%
    select(c("Objs","Sensitivity"))
  print(obj)
}

printSensitivityObj(model)
printSensitivityRHS(model)

```

# Questions about LP

1.	Can an LP model have more than one optimal solution. Is it possible for an LP model to have exactly two optimal solutions? Why or why not?

Some LP models can have more than one solution. They can have alternate optimal solution.In fact, there can be more than one feasible point that maximized orminimized the value of the objective function.
In many LP models we can see that the final level curve intersect the feasible region along an edge of the feasible region, and not only in a singular point. 
All the point of this intersection produce the same objective target value. All of this points are alternate optimal solutions. For this reason for an LP model is not possible to have exactly two optimal solution. An LP model can have exactly one optimal solution because this point is the intersection between the final level curve and the feasible region.

2. Are the following objective functions for an LP model equivalent? That is, if they are both used, one at a time, to solve a problem with exactly the same constraints, will the optimal values for $x_1$ and $x_2$ be the same in both cases? Why or why not?

Hi, welcome to JustAnswer!

Yes, provided an optimal solution exists, the optimal (x1, x2) will 

remain the same for both maximization and minimization problem.

This is because the the feasible region is decided by the constraints 

and as long as the constraints remain the same so does the feasible 

region, and the corner point (x1, x2) that maximizes or minimizes the 

objective function.

$$
\max 2x_1+3x_2\\
\min -2x_1-3x_2
$$

3. Which of the following constraints are not linear or cannot be included as a constraint in a linear programming problem? 


a. $2 x_1+x_2-3 x_3 \geq 50$   

b. $2 x_1+\sqrt{x_2} \geq 60$

c. $4 x_1- \frac{1}{2} x_2=75$

d. $\frac{3 x_1+2 x_2x-3 x_3}{x_1+x_2+x_3} \leq 0.9$

e. $3 x_1^2+7 x_2 \leq 45$



A linear constraint is a mathematical expression where the decision variables are added or subtracted and the result of the expression is grather than or equal, less tahn or equal, or equal to a the right hand side value.


a. The equation is linear. 
b. The equation in non-linear
c. The equation is linear
d. The equation is non-linear
e. The equation is not linear








